Clouds and Radiation in a mock-Walker Circulation

### 2019-09-20 Silvers_WalkerCell.tex
changed eps file name
substantial editing of the text for Silvers_WalkerCell.tex

@@

To develop the model configuration used for these experiments we started with the same code base as that of the AM4.0 
model (Zhao et al. a,b).  This includes the GFDL finite-volume cubed-sphere dynamical core FV3 (Harris and Lin, 2013) 
which can solve either the hydrostatic primitive equations or the nonhydrostatic fully compressible Euler equations.  Taking 
the AM4.0 physics we use the interactive radiation, deep and shallow convection, large-scale clouds, and boundary layer 
scheme as described in Zhao et al. a,b and the references therein.  The prognostic moisture variables are the specific 
humidity (q), liquid (ql) and frozen water (qi), and cloud fraction.  The top of the model domain is at 1 hPa, with 33 vertical 
levels and a sponge layer extending downward to 8 hPa.  The kilometer of atmosphere just above the surface is resolved by 
8 model levels.  The changes made to the default AM4.0 physics configuration are the following.  The cloud-aerosol 
interactions were turned off to focus on the interaction between clouds, radiation, and the circulation.  The gravity wave drag 
parameterization was turned off 
%(do_cg_drag set to .f.; Alexander-Dunkerton gravity wave drag) 
in order to reduce large oscillations in the horizontal wind field in the top of the model domain.  The convection, radiation, 
large-scale cloud, microphysics, and turbulence parameterization all remain the same as in AM4.0.  
Thus for the experiments with convective parameterization (grid-spacing of 25km and 100km), the physics are very 
similar to those of AM4.0.  This configuration can also be compared to the simulations described in 
Popp and Silvers, 2017, which used a developmental version of AM4 with almost the same changes to the physics
schemes.
For our simulations using grid spacing of 1km, 2km, and 25km (in select simulations), the double plume (both shallow and deep convection) convective parameterization was 
turned off and the time step of integration increased.

@@

All simulations use a nonhydrostatic dynamical core, with prescribed SSTs and a doubly periodic domain which is elongated in the zonal 
direction allowing for three dimensional simulations but with a reduced computational cost relative to the global domain which is usually 
used with this model.  The SST is prescribed as a Gaussian function which is 4K warmer in the center (301K/27.85C) of the domain then at the edges (297K/23.85C).  
Thus far the GCM runs with 25km resolution have been run for 5 year.  
The 1km and 2km CRM were run for 6 months.

The simulations with a grid-spacing of 1km and 2km are referred to as cloud resolving simulations.  For our purposes this simply means
that both deep and shallow convective parameterizations are turned off, and that the threshold of grid-cell mean relative humidity 
which triggers new clouds is changed from the default value of 0.8 to 1.0.   The Tiedtke (1993) parameterization scheme for large-scale 
clouds was designed to be used with GCMs having a coarse grid-spacing.  However, there is not a fundamental problem that we are 
aware of in using the Tiedtke scheme for large-scale clouds in a model with 1km grid-spacing.   The advantage is retaining the 
identical cloud scheme as is used in the parent GCM; the disadvantage is the greatly increased complexity of the cloud computations 
relative to many other cloud resolving models.

One tool that has been commonly used to infer the influence of clouds on a model simulation is to make the clouds invisible to the 
radiation.  This can tell us the extent to which the specific location of the clouds matters to the radiative interaction with the atmospheric state.   This method was originally pioneered by Slingo and Slingo, 1988 and Randall et al., 1989.  More recently, it has been
implemented as part of the CFMIP series of experiments (Stevens et al. 2012).
In the AM4 code, this is done separately for the longwave and shortwave radiation.  In this study we compare control experiments, 
in which clouds and radiation are fully interactive with experiments in which clouds are
invisible to the longwave radiation.  These experiments are referred to as Longwave Cloud Radiative Effect Off (LWCRE off).  
Other than the invisible clouds, the model is the same.  Both longwave and shortwave radiation are present and interactive with 
the atmospheric state, the clouds still interact with the shortwave radiation, and they still precipitate.   
In general, turning off the longwave cloud radiative effect will have a large impact on the surface budget of a coupled model.  
However, because there is no land in our simulations and the SST is held fixed, the energetics of our experiments are not as
strongly effected as might be expected.  Experiments with only a water surface at the lower boundary and fixed SST are the
ideal configuration to utilize the LWCRE off experiments.

\section{Results}

The results of our experiments are organized as follows.  First, the general structure and characteristics of the steady state solutions
will be described.   Then, the influence of parameterized convection and the longwave cloud radiative effect on the overturning 
circulation will be shown.  We will then briefly discuss how the circulation changes as a function of domain size.  The last part of the
results will describe and contrast the Walker circulation in a GCM-like configuration and a CRM-like configuration.

The mock-Walker Circulation that emerges from these simulations is a deep overturning circulation.  There are two circulation cells, 
one below $500 \rm{hPa}$  and one above.  Deep-convective activity dominates the region over the warm pool, shallow convection is 
common over a wider range of SSTs, and for the most part stratocumulus 
clouds are absent.  When the coupling between the circulation and clouds is broken, by turning off the longwave CRE, the atmospheric 
state is much more symmetric about the maximum SST.

@@

@@

\subsection{The Asymmetric Influence of Convective Parameterization and the organizing tendency of LWCRE}

@@

To illustrate how the double plume convective parameterization influences the atmospheric state we compare the P25 and
E25 simulations in which the only difference is that all convective parameterization is turned off in E25 (refer to a figure with
profiles for E25 and P25).  The P25 experiment represents a GCM-like configuration and compared to E25 has a stronger 
overturning circulation, higher humidity in the deep convective region, and lower humidity in the subsidence region over the 
cooler SST (Figure \ref{fig:rh_psi_P25vsE25}).  Averaged over the full domain, the parameterized convection leads to a 
dryer atmosphere with less 
condensate (both liquid and ice) and with weaker radiative cooling and condensational heating (below 800 hPa).  However, in 
the updraft region the convective parameterization results in more moisture aloft.

@@

@@

\subsection{Dependence on Domain Size}

% PDF for location of precip maximum of precipitation for experiments at both 25km and 100km resolutions 
% LW CRE off
\begin{figure}
  \centering
      %\includegraphics[width=0.96\columnwidth]{walkerfigs/hist_4pan_25vs100_lwcreoff.eps}
      \includegraphics[width=0.6\columnwidth]{walkerfigs/hist_4pan_25vs100_lwcre_draft.eps}
  \caption{Normalized probability distributions of the location of the maximum value of precipitation for P25 and P100
  experiments .   Panels on the right (left) show experiments on the domain with a long dimension of 4000 km 
  (16000 km).  Black distributions are for experiments with the LW CRE on and green distributions for experiments 
  with LW CRE off.  The black curves show a Gaussian function with a standard deviation computed for 
  each experiment and an expected location in the center of the domain where the SST is warmest.}
\end{figure}

When designing numerical experiments the size of the domain and the grid-spacing that determines model resolution are two of the most 
fundamental choices that must be made.   One the goals of this study was to explore the results of changing grid-spacing over a wide 
range of values.  To simplify the analysis we have in most cases chosen to keep the 
long horizontal dimension fixed at $4000\, \rm{km}$.  However, previous studies have documented sensitivities to domain size.  With this 
in mind several different domain shapes and sizes have been explored, but always keeping the $4000\, \rm{km}$ long-dimension as the 
primary domain for the analysis.  
Examining the evolution in time of the precipitation field reveals important characteristics of these simulations, and 
illustrates how much the distribution can vary as a function of domain size, parameterization of convection, and the effect of the 
longwave radiation due to clouds.  At all resolutions the Hovmoller plots show that the longwave CRE acts to concentrate the 
precipitation over a smaller geographic extent.  
Shown in Figure {\ref{fig:domdep}} are Hovmoller plots of precipitation after 
averaging along the short horizontal dimension.  The four panels show simulations with a grid-
spacing of $25\, \rm{km}$ on a grid with a horizontal width of $4000\, \rm{km}$ (far right) and $16000\, \rm{km}$ (middle right), and 
simulations with a grid-spacing of $100\, \rm{km}$ on a grid with a horizontal width of $4000\, \rm{km}$ (middle left) and 
$16000\, \rm{km}$ (far left).  Figure  {\ref{fig:domdep_lwoff}} shows the equivalent simulations with LW CRE off.    
The structure of the precipitation depends strongly on domain size and shows less dependence on grid-spacing.

Smaller domain lengths were found to have a more focused ascent region and larger precipitation rates in Bretherton et al., 2006.
They also found less low-level clouds over the colder SSTs in the smaller domains, relative to a control domain size that was 4x
larger.   We have found larger precipitation rates for the larger domains.  Does our larger domain ascent region change?  Does 
Brethertons SST warm patch size increase with domain size?

% Fraction of Precip due to large-scale precip for large/small experiments with a 25km resolution
\begin{figure}
  \centering
      \includegraphics[width=0.8\columnwidth]{walkerfigs/Conv_vs_LS_hist.eps}
  \caption{Fraction of precipitation that is due to the large-scale(blue) cloud scheme relative to the total 
  precipitation (large-scale plus convective).}
  \label{fig:conv_vs_ls}
\end{figure}

\subsection{The mock-Walker Circulation Dependence on Grid-Spacing: A transition from a 
General Circulation Model to a Cloud Resolving Model}

It is clear that the overturning circulation in the tropical Pacific encompasses dynamical motions at a host of scales and that the 
different cloud types interact with the circulation on scales ranging from meters to thousands of km's.  
Our primary motivation for using the GFDL AM4 on this doubly-periodic domain is that the dynamical core of the model can 
easily be used over a wide range of grid-spacings and can be used to solve either the hydrostatic or nonhydrostatic 
governing equations.  In this section we focus on simulations with grid-spacings that range between 1km and 100km.   
A glance at Figure \ref{fig:hov_4mods_6mn} shows notable differences in the precipitation structure that results from the 
overturning circulation at different resolutions.  Shown are the first 180 days of the P25, E25, E2, and E1 simulations.  As the 
grid-spacing decreases towards $1 \, \rm{km}$ the distribution of precipitation becomes more regular and consistently 
centered over the SST maxima in the middle of the domain.  Both the P25 and E25 simulations show more variability at later 
times, compared to these first 180 days.  Due to the large expense of the E2 and E1 simulations they were run for 
180 days.  The simulations with explicit convection at resolutions typical of cloud-resolving models (E2, E1) show little 
aversion to the precipitation maximum occurring over the maximum in SST and relative to the P25 and E25 simulations 
they are able to maintain a smoother distribution of  precipitation values over a broader range of SST values.

The steady state relative humidity and overturning circulation show an interesting combination of similarities and differences
among the experiments with varying grid-spacing.  All experiments have a multicell vertical structure (consistent with 
Bretherton et al. 2006), in contrast to the single cell that results from simple theoretical models such as the SQTCM.   All 
show a similarly dry mid-tropospheric region where subsidence dominates.  However, the multicelled structure is quite irregular
in the P25 experiment, and a small third cell has developed in the boundary layer of the 1km experiment.  The mid-troposphere 
in the region with deep-convection is more than $40\%$ dryer than it is for the 2km and 1km experiments.

Despite a fairly regular distribution of precipitation around the SST maximum for experiments with increasing resolution, the 
surface enthalpy flux reveals large differences in the symmetry.  Over the SST maximum, E25 has a surface enthalpy flux that is 
60 W/m2 larger than that of the E1 experiment, and the E1 experiment has a very irregular pattern of enthalpy flux in the middle
half of the domain.  It appears that these differences in magnitude and regularity are due to the differing wind speeds at the 
lowest atmospheric level.  It is also work noting that the E1 experiment has a stronger domain mean u wind shear compared 
with the E2 and E25 experiments.  Over the warm patch of SST, the surface enthalpy flux is decreased by the LWCRE for E25, 
E2, and E1 despite these experiments having stronger low-level winds.  This decreased enthalpy flux also corresponds to 
an increased amount of specific humidity in the lowest atmospheric model level.        
%%%%%
% section 3 of results

% Hovmoller diagram of precipitation for the first 6 months of the 25km, 2km, and 1km experiments
% LW CRE on
\begin{figure}
  \includegraphics[width=0.96\columnwidth]{walkerfigs/precip_4mods_6mn.eps}
  \caption{Evolution of precipitation through the first 6 months of simulation for (left to right) the 25km control case, 25km case with no 
  parameterized convection, 2km control, and 1km control.  Shading indicates mm/day.  The horizontal axis shows the entire 
  longitudinal domain of 4000 km with the center of the domain having a prescribed SST of 301 K and the edges 297 K.  
  Shown is the meridional average. }
  \label{fig:hov_4mods_6mn}
\end{figure}

@@

Because of the close connection between the domain mean value of precipitation, atmospheric condensational heating, and 
the total radiative cooling the time evolution of the precipitation serves as a useful measure of wether a model has 
reached statistical equilibrium.  Figure \ref{fig:precip_dom_mn} demonstrates that after about 30 days, the P25, E2, and E1
simulations begin to oscillate about a mean state that is roughly constant over periods greater than about 30 days. 
The large oscillations in domain mean precipitation are similar to those noted in previous studies 
(Silvers et al., 2016; Patrizio and Randall, 2019).   The experiments with a higher resolution have a lower value of domain
mean precipitation, with the E2 simulation having the lowest (3.1 mm/day).  This can be understood as a consequence
of the differences in upper level cloud fraction.   Figure \ref{fig:cf_tdtlw} shows E2 to have the largest (about 17\%) upper level mean cloud
fraction in the subsidence region, with the E1 experiment having the next largest cloud fraction (10\%), followed by E25, P25, and 
P100 (5\%).   Less upper level clouds will allow more radiative cooling of the atmosphere and require a correspondingly larger
amount of condensational heating and precipitation.

@@

The domain mean condensate is an important measure of the steady state reached by these experiments.  The vertical 
distribution of condensate is closely related to the distribution of clouds, it influences the distribution of energy throughout
the atmosphere as well as influencing the outgoing longwave radiation.  It can also indicate the strength of convection 
and vertical mass transport.     Figure \ref{fig:TotCond} shows the domain mean condensate for P25, E25, E2, and E1
(solid lines) and the corresponding experiments with the LWCRE off (dashed lines).  There is a fundamental difference 
between the GCM-like P25/E25 experiments and the CRM-like E2/E1 experiments with the former having much 
higher values of low-level liquid condensate/cloud and the later having much higher values up upper level ice condensate.  
It is also interesting to note that the LWCRE results in less condensate above 700 hPa for all experiments, and 
more condensate below 700 hPa for the P25/E25 experiments.  The two-dimensional structure of circulation and humidity
shows that the amount of condensate aloft in the deep-convective region increases with resolution while the relative
humidity in the subsidence regions decreases with resolution.  This is consistent with an overturning circulation that 
strengthens as the resolution increases.   This dependence of circulation strength and moisture on resolution is especially 
apparent when the LWCRE is off (Figure \ref{fig:rh_psi_P25E2E1_lwoff}).  When the clouds and radiation are allowed 
to directly interact with each the E2 and E1 simulations appear much more similar (Figure \ref{fig:rh_psi_P25E2E1}).

The E25 and P25 experiments have much less condensate aloft and a lower relative humidity in the mid-troposphere 
(can we use this to understand the differences in LWCRE?).

@@

Superposing the the circulation and relative humidity illustrates the subsidence driven 
drying and the strong moistening that results from ascent over the region of maximum latent heat flux.

@@

\textit{probably insert figure from slides 20 and 21 of IUGG talk.}
We can see clearly that the interactions between longwave radiation and clouds act to dry the troposphere in the regions of subsidence 
more than the cases when the longwave CRE is not active.  This is especially true for the simulations with a grid-spacing of 1 km 
and 2 km.  While the mid-tropospheric profiles of diabatic heating and cooling are similar between the LW CRE on/off simulations, 
there is a very strong response below about 850 hPa for the 25km model.  At that grid-spacing, with the LW CRE on, significant 
low-level clouds are formed, leading to a radiative cooling of almost 10K/day, and a radiative cooling of around -7K/day.   However, 
when the LW CRE is off, there is no appreciable difference in the low-level diabatic profiles among the models with 3 grid-spacings.   It 
is also clearly seen that the LW CRE decreases the upper-level ice-condensate for models of all grid-spacings while dramatically 
increasing the low-level liquid condensate.

@@

>>>>

The results of our experiments are organized as follows.  First, the general structure and characteristics of the steady state solutions
will be described.   Then, the influence of parameterized convection and the longwave cloud radiative effect on the overturning 
circulation will be shown.  We will then briefly discuss how the circulation changes as a function of domain size.  The last part of the
results will describe and contrast the Walker circulation in a GCM-like configuration and a CRM-like configuration as the 
resolution changes.

@@

To develop the model configuration used for these experiments we started with the same code base as that of the 
recently developed atmospheric global climate model AM4.0 
(Zhao et al. a,b).  This includes the GFDL finite-volume cubed-sphere dynamical core FV3 (Harris and Lin, 2013) 
which can solve either the hydrostatic primitive equations or the nonhydrostatic fully compressible Euler equations.  
The default AM4.0 physics we use includes interactive radiation, parameterized deep- and shallow-convection, 
a large-scale cloud scheme, and a boundary layer 
scheme as described in Zhao et al. a,b and the references therein.  The prognostic moisture variables are the specific 
humidity (q), liquid (ql) and frozen water (qi), and cloud fraction.  The top of the model domain is at 1 hPa, with 33 vertical 
levels and a sponge layer extending downward to 8 hPa.  The kilometer of atmosphere just above the surface is resolved by 
8 model levels.  The changes made to the default AM4.0 physics configuration are the following.  The cloud-aerosol 
interactions were turned off to focus on the interaction between clouds, radiation, and the circulation.  The gravity wave drag 
parameterization was turned off 
%(do_cg_drag set to .f.; Alexander-Dunkerton gravity wave drag) 
in order to reduce large oscillations which developed in the horizontal wind field near the top of the model domain.  
The convection, radiation, large-scale cloud, microphysics, and turbulence parameterizations all remain the same 
as in AM4.0.   Thus for the experiments with convective parameterization (grid-spacing of 25km and 100km), the 
physics are very similar to those of AM4.0.  This configuration can also be compared to the simulations described in 
Popp and Silvers, 2017, which used a developmental version of AM4 with almost the same changes to the physics
schemes.

@@

All simulations use a nonhydrostatic dynamical core, with prescribed SSTs and a doubly periodic domain which is elongated in the zonal 
direction allowing for three dimensional simulations but with a reduced computational cost relative to the default global domain.  The SST is prescribed as a Gaussian function which is 4K warmer in the center (301K/27.85C) of the domain then 
at the edges (297K/23.85C).  The GCM runs with 25km and 100km grid-spacing have been run for 5 years while the 
1km and 2km experiments were run for 6 months.  Experiments with parameterized convection will be labelled with a P prefix, followed
by a number indicating the grid-spacing in kms while the experiments with explicit convection and no parameterized convection will be 
labelled with an E prefix, followed by the appropriate number.  Thus P25 refers to an experiment with parameterized convection using a 
grid-spacing of 25km, while E25 refers to an experiment that is identical to P25 except that the convective parameterization is turned off.

To illustrate the strong dependence of our results on domain size, as well as the fundamental role that the LW CRE plays in GCMs
we utilize fully parameterized experiments with grid-spacing of 25km (P25) and 100km (P100) on domains with a long dimension 
of 4000 km (`small') and
16000 km (`large').  To explore the mock-Walker circulation in the context of both a GCM and a CRM we utilize 
comparisons of the experiments with grid-spacings of 25km (E25), 2km (E2), and 1km (E1) all on a domain with the same long dimension 
of 4000 km.  The experiments with a grid-spacing of 25km serve as a link between the GCM-like configuration and the CRM-like 
configuration.  Domains with dimensions of 16000 km are too costly for the 1km and 2km experiments.

The E1 and E2 simulations are in many ways similar to the configuration of so-called cloud resolving models.  
In particular both the deep and shallow convective parameterizations are turned off, and the threshold of grid-cell mean relative humidity 
which triggers new clouds is changed from the default value of 0.8 to 1.0.  While a grid spacing of 1 or 2km is clearly not small enough
to resolve all clouds, it is small enough to resolve cloud-systems and medium to larger sized clouds.  The primary 
difference between the CRM simulations presented in this paper and those of more established CRMs such as SAM is that our model
has a much coarser vertical resolution and a prognostic large-scale cloud scheme inherited from the GCM.  The Tiedtke (1993) 
parameterization scheme for large-scale 
clouds was designed to be used with GCMs having a coarse grid-spacing.  However, there is not a fundamental problem that we are 
aware of in using the Tiedtke scheme for large-scale clouds in a model with 1km grid-spacing.   The advantage is retaining the 
identical cloud scheme as is used in the parent GCM; the disadvantage is the greatly increased complexity of the cloud computations 
relative to many other cloud resolving models.

One tool that has been commonly used to infer the influence of clouds on a model simulation is to make the clouds invisible to the 
radiation.  This can tell us the extent to which the specific location of the clouds matters to the radiative interaction with the 
atmospheric state.   This method was originally pioneered by Slingo and Slingo, 1988 and Randall et al., 1989.  More recently, it has been
implemented as part of the CFMIP series of experiments (Stevens et al. 2012).
In the AM4 code, this is done separately for the longwave and shortwave radiation.  In this study we compare control experiments, 
in which clouds and radiation are fully interactive with experiments in which clouds are invisible to the longwave radiation.  These 
experiments are referred to as Longwave Cloud Radiative Effect Off (LWCRE off).  
For the LWCRE off experiments, both the longwave and shortwave radiation are present and interact with 
the atmospheric state, the clouds still interact with the shortwave radiation, and they still precipitate.   
Normally, turning off the longwave cloud radiative effect would have a large impact on the surface budget of a coupled model.  
However, because there is no land in our simulations and the SST is held fixed, the energetics of our experiments are not as
strongly effected as might be expected.  For this reason, experiments with only a water surface at the lower boundary and 
fixed SST are the ideal configuration to utilize the LWCRE off experiments as a way of diagnosing how clouds interact with 
the atmospheric state.

\section{The organizing influence of LWCRE and the asymmetric results of parameterized convection.}

The mock-Walker Circulation that emerges from these simulations is shown in Figures \ref{fig:rh_psi_P25vsE25} and 
\ref{fig:precip_vertvel} to be characterized by a strong overturning circulation with precipitation focused over the warmer 
SSTs and a humid boundary layer across the full length of the domain.   For the GCM-like configuration (P25, left panels 
of Figure \ref{fig:rh_psi_P25vsE25}) with fully parameterized convection and with the longwave CRE turned on, there is 
a strong overturning circulation that results in a humid, condensate loaded troposphere near the middle of the domain 
and a dry ($< 10\% RH$) subsidence region with almost no condensate above $800 \rm{hPa}$.    
To illustrate how (how what?  the double-plume convective parameterization influences the atmospheric state) 
we compare the P25 experiment with analogous experiments in which the longwave CRE is off (P25 LWCRE off, middle 
panels of Figure \ref{fig:rh_psi_P25vsE25}) and in which the convection is made explicit by turning off the convective 
parameterization (E25, right panels of Figure \ref{fig:rh_psi_P25vsE25}).  
The mass streamfunction and vertical velocity both show the P25 experiment has a stronger overturning circulation 
than either E25 or P25 LWCRE off.  P25 also has higher humidity in the deep convective region, and lower humidity in 
the subsidence region over the 
cooler SST (Figure \ref{fig:rh_psi_P25vsE25}) than either E25 or P25 LWCRE off.  Averaged over the full domain, 
the parameterized convection leads to a dryer atmosphere with less 
condensate (both liquid and ice) and with weaker radiative cooling and condensational heating (below 800 hPa).  However, in 
the updraft region the convective parameterization results in more moisture aloft.

There are circulation cells below $500 \rm{hPa}$  and above $500 \rm{hPa}$.  Deep-convective activity dominates the 
region over the warm pool, shallow convection is common over a wider range of SSTs, and for the most part stratocumulus 
clouds are absent.  When the coupling between the circulation and clouds is broken, by turning off the longwave CRE, the atmospheric 
state is much more symmetric about the maximum SST.

@@

The abnormality, or asymmetry, of the P25 simulation was the first surprising feature noticed by the authors.
This asymmetry is also present on larger domains and at a resolution of $100 \, \rm{km}$.  For some of the other cases observed, such
as P100 (on the small domain), the asymmetry can be described as an aversion of the precipitation maxima to reside over the 
SST maximum while in other cases, such as the P25 case shown in \ref{fig:rh_psi_P25vsE25}, a strong ($1 \, \rm{m s^-1}$) domain
mean shear develops which shifts the precipitation and circulation off center for years at a time.  This appears to be not entirely a 
result of the convective parameterization, but due to an interaction between the convective parameterization and the LW CRE.

@@

%\subsection{The Asymmetric Influence of Convective Parameterization and the organizing tendency of LWCRE}

@@

@@

\subsection{The Influence of Domain Size on the LWCRE and the LS Parameterized Precipitation}

%% PDF for location of precip maximum of precipitation for experiments at both 25km and 100km resolutions 
%% LW CRE off
%\begin{figure}
%  \centering
%      %\includegraphics[width=0.96\columnwidth]{walkerfigs/hist_4pan_25vs100_lwcreoff.eps}
%      \includegraphics[width=0.6\columnwidth]{walkerfigs/hist_4pan_25vs100_lwcre_draft.eps}
%  \caption{Normalized probability distributions of the location of the maximum value of precipitation for P25 and P100
%  experiments .   Panels on the right (left) show experiments on the domain with a long dimension of 4000 km 
%  (16000 km).  Black distributions are for experiments with the LW CRE on and green distributions for experiments 
%  with LW CRE off.  The black curves show a Gaussian function with a standard deviation computed for 
%  each experiment and an expected location in the center of the domain where the SST is warmest.}
%\end{figure}
%

When designing numerical experiments the size of the domain and the grid-spacing that determines model resolution are two of the most 
fundamental choices that must be made.   One the goals of this study was to explore the results of changing grid-spacing over a wide 
range of values.  To simplify the analysis we have in most cases chosen to keep the 
long horizontal dimension fixed at $4000\, \rm{km}$.  However, previous studies (e.g. Bretherton et al. 2006) have 
documented sensitivities to domain size.   The analysis of the previous, and of the next section focus on results from experiments using
a domain with a long domain length of $4000 \, \rm{km}$.  However we do find sensitivity to the domain size and thought it to be
sufficiently interesting to report.

Examining the evolution in time of the precipitation field reveals important characteristics of these simulations, and 
illustrates how much the distribution can vary as a function of domain size, parameterization of convection, and the effect of the 
longwave radiation due to clouds. 
Shown in Figures {\ref{fig:domdep}} and {\ref{fig:domdep_lwoff}} are Hovmoller plots of precipitation after 
averaging along the short horizontal dimension.  The four panels show simulations with a grid-
spacing of $25\, \rm{km}$ on a grid with a horizontal width of $4000\, \rm{km}$ (far right) and $16000\, \rm{km}$ (middle right), and 
simulations with a grid-spacing of $100\, \rm{km}$ on a grid with a horizontal width of $4000\, \rm{km}$ (middle left) and 
$16000\, \rm{km}$ (far left).  Figure  {\ref{fig:domdep_lwoff}} shows the equivalent simulations with LW CRE off.    
 At all resolutions the Hovmoller plots show that the longwave CRE acts to concentrate the 
precipitation over a smaller geographic extent.   This is consistent with the idea that CREs strengthen an overturning circulation and
narrow the region of deep convection (see Popp and Silvers, 2017, HH, Dixit et al. 2019.  etc.).
The structure of the precipitation depends strongly on domain size and shows less dependence on grid-spacing.
On the large domains, the difference between experiments with and without LW CRE is extreme.  In contrast to the control 
experiments in Figure {\ref{fig:domdep}} which all show a narrow region of strong precipitation which meanders within 
500 km of the SST maximum at the center of the domain,  
the experiments without the LW CRE basically have no region with consistently strong precipitaiton.  Instead, 
smaller cells or lines of precipitation develop within an area that is roughly 8000 km wide.

Smaller domain lengths were found to have a more focused ascent region and larger precipitation rates in Bretherton et al., 2006.
They also found less low-level clouds over the colder SSTs in the smaller domains, relative to a control domain size that was 4x
larger.   We have found larger precipitation rates for the larger domains.  Does our larger domain ascent region change?

% Fraction of Precip due to large-scale precip for large/small experiments with a 25km resolution
\begin{figure}
  \centering
      \includegraphics[width=0.8\columnwidth]{walkerfigs/Conv_vs_LS_hist.eps}
  \caption{Precipitation that is due to the large-scale(blue) cloud scheme and to the convective parameterization (black).}
  \label{fig:conv_vs_ls}
\end{figure}

@@

\section{The mock-Walker Circulation Dependence on Grid-Spacing: A transition from a 
General Circulation Model to a Cloud Resolving Model}

It is clear that the overturning circulation in the tropical Pacific encompasses dynamical motions at a host of scales and that the 
different cloud types interact with the circulation on scales ranging from meters to thousands of km's.  
The tropical Pacific also serves as an ideal location to study interactions between clouds and the 
circulation because it includes strong overturning circulations, abundant shallow cumulus, congestus, and cumulonimbus 
clouds (Johnson et al. 1999) as well as stratocumulus cloud decks along the eastern extremities of the basin.
Can we hope to simulate all of this in an idealized modeling context?    
Our initial motivation for using the GFDL AM4 model on this doubly-periodic domain was to simulate this region 
%that is rich in phenomena 
in the context of both a GCM and a CRM in the hopes of learning something from the high resolution 
cases that will help us to improve the development of the GCM.    
The dynamical core of the current generation GFDL models is ideally suited to this task because it can easily 
be used over a wide range of resolutions and can solve either the hydrostatic or nonhydrostatic 
governing equations.  In this section we focus on simulations with grid-spacing of 1km, 2km, and 25km.   
A glance at Figure \ref{fig:hov_4mods_6mn} shows notable differences in the precipitation structure that result from the 
overturning circulation at different resolutions.  Shown are 180 days precipitation from the P25, E25, E2, and E1 simulations.  As the 
resolution increases towards $1 \, \rm{km}$ the distribution of precipitation becomes more regular and consistently 
centered over the SST maxima in the middle of the domain.  Both the P25 and E25 simulations show more variability at later 
times, compared to these first 180 days.  The simulations with explicit convection at resolutions typical of cloud-resolving 
models (E2, E1) show little aversion to the precipitation maximum occurring over the maximum in SST and relative to 
the P100, P25 and E25 simulations 
they are able to maintain a smoother distribution of  precipitation values over a broader range of SST values.

\begin{figure}
  \includegraphics[width=0.96\columnwidth]{walkerfigs/precip_4mods_6mn.eps}
  \caption{Evolution of precipitation through the first 6 months of simulation for (left to right) the 25km control case (P25), 
  25km case with no parameterized convection (E25), 2km control (E2), and 1km control (E1).  
  Shading indicates mm/day.  The horizontal axis shows the entire 
  longitudinal domain of 4000 km with the center of the domain having a prescribed SST of 301 K and the edges 297 K.  
  Shown is the meridional average. }
  \label{fig:hov_4mods_6mn}
\end{figure}

Because of the close connection between the domain mean value of precipitation, atmospheric condensational heating, and 
the total radiative cooling the time evolution of the precipitation serves as a useful measure of wether a model has 
reached statistical equilibrium.  Figure \ref{fig:precip_dom_mn} demonstrates that after about 30 days, the P25, E2, and E1
simulations begin to oscillate about a mean state that is roughly constant over periods greater than about 30 days. 
The large oscillations in domain mean precipitation are similar to those noted in previous studies 
(Silvers et al., 2016; Patrizio and Randall, 2019).   The experiments with a higher resolution have a lower value of domain
mean precipitation, with the E2 simulation having the lowest (3.1 mm/day).  This can be understood as a consequence
of the differences in upper level cloud fraction and will be further discussed.

Superposing the circulation and relative humidity illustrates the subsidence driven 
drying and the strong moistening that results from ascent over the region of maximum latent heat flux.

The steady state relative humidity and overturning circulation show an interesting combination of similarities and differences
among the experiments with varying resolutions.  All experiments have a multicell vertical structure (consistent with 
Bretherton et al. 2006), in contrast to the single cell that results from simple theoretical models such as the SQTCM.   All 
show a similarly dry mid-tropospheric region where subsidence dominates.  However, the multicelled structure is quite irregular
in the P25 experiment, and a small third cell has developed in the boundary layer of the 1km experiment.  The mid-troposphere 
of the P25 experiment is more than $40\%$ dryer in the region with deep-convection than for the 2km and 1km experiments.

Despite a fairly regular distribution of precipitation around the SST maximum for experiments with increasing resolution, the 
surface enthalpy flux reveals large differences in the symmetry.  Over the SST maximum, E25 has a surface enthalpy flux that is 
60 W/m2 larger than that of the E1 experiment, and the E1 experiment has a very irregular pattern of enthalpy flux in the middle
half of the domain.  It appears that these differences in magnitude and regularity are due to the differing wind speeds at the 
lowest atmospheric level.  It is also worth noting that the E1 experiment has a stronger domain mean u wind shear compared 
with the E2 and E25 experiments.  Over the warm patch of SST, the surface enthalpy flux is decreased by the LWCRE for E25, 
E2, and E1 despite these experiments having stronger low-level winds.  This decreased enthalpy flux also corresponds to 
an increased amount of specific humidity in the lowest atmospheric model level.        
%%%%%
% section 3 of results

% Hovmoller diagram of precipitation for the first 6 months of the 25km, 2km, and 1km experiments
% LW CRE on

@@

Figure \ref{fig:cf_tdtlw} shows E2 to have the largest (about 17\%) upper level mean cloud
fraction in the subsidence region, with the E1 experiment having the next largest cloud fraction (10\%), followed by E25, P25, and 
P100 (5\%).   Less upper level clouds will allow more radiative cooling of the atmosphere and require a correspondingly larger
amount of condensational heating and precipitation.

@@

The domain mean condensate is an important measure of the steady state reached by these experiments.  The vertical 
distribution of condensate is closely related to the distribution of clouds, it influences the distribution of energy throughout
the atmosphere as well as influencing the outgoing longwave radiation.  It can also indicate the strength of convection 
and vertical mass transport.     Figure \ref{fig:TotCond} shows the domain mean condensate for P25, E25, E2, and E1
(solid lines) and the corresponding experiments with the LWCRE off (dashed lines).  There is a significant difference 
between the GCM-like P25/E25 experiments and the CRM-like E2/E1 experiments with the former having much 
higher values of low-level liquid condensate/cloud and the later having much higher values of upper level ice 
condensate.  The LWCRE off experiments influence the low-levels of the GCM-like experiments, and the upper-level
regions of the CRM-like experiments.     
It is also interesting to note that the LWCRE results in less condensate above 700 hPa for all experiments, and 
more condensate below 700 hPa for the P25/E25 experiments.  The two-dimensional structure of circulation and humidity
shows that the amount of condensate aloft in the deep-convective region increases with resolution while the relative
humidity in the subsidence regions decreases with resolution.  This is consistent with an overturning circulation that 
strengthens as the resolution increases.  When the clouds and radiation are allowed 
to directly interact with each the E2 and E1 simulations appear much more similar (Figure \ref{fig:rh_psi_P25E2E1}).   
This dependence of circulation strength and moisture on resolution is especially 
apparent when the LWCRE is off (Figure \ref{fig:rh_psi_P25E2E1_lwoff}).

The E25 and P25 experiments have much less condensate aloft and a lower relative humidity in the mid-troposphere
where the differences in strength of the overturning circulation are most pronounced. 
(can we use this to understand the differences in LWCRE?).

@@

@@

We can see clearly that the interactions between longwave radiation and clouds act to dry the troposphere in the regions of subsidence 
more than the cases when the longwave CRE is not active.  This is especially true for the simulations with a grid-spacing of 1 km 
and 2 km.  While the mid-tropospheric profiles of diabatic heating and cooling are similar between the LW CRE on/off simulations, 
there is a very strong response below about 850 hPa for the 25km model.  At that grid-spacing, with the LW CRE on, significant 
low-level clouds are formed, leading to a radiative cooling of almost 10K/day, and a radiative cooling of around -7K/day.   However, 
when the LW CRE is off, there is no appreciable difference in the low-level diabatic profiles among the models with 3 grid-spacings.   It 
is also clearly seen that the LW CRE decreases the upper-level ice-condensate for models of all grid-spacings while dramatically 
increasing the low-level liquid condensate.

@@

References:
